---
title: "manhattan_final"
author: "Ruilin"
date: "2023-06-19"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Loading of data, removal of incomplete entries and dividing into
train and test data

```{r}
#install.packages("gridExtra")
library(gridExtra)
set.seed(302)
data=read.csv("manhattan.csv")
attach(data)
data1= na.omit(data)
rows <- sample(1:3539, 2800, replace=FALSE) 
train<- data1[rows,]
test=data1[-rows,]
```

Summary of variables of the full model

```{r}


summary(train[,c(3,4,5,6,2)])



summary(test[,c(3,4,5,6,2)])


```

Original EDA, histgrams,boxplots and scatterplots
```{r}


attach(train)
par(mfrow=c(3,2))
hist(rent, breaks=10, main="Rent",col="#1793d1")
hist(bedrooms,breaks=10,xlab = "number of bedrooms", main="Number of Bedrooms")
hist(bathrooms, breaks=10,xlab = "number of bathrooms", main="Number of Bathrooms")
hist(size_sqft, breaks=20,xlab = "size", main="Size in sqft")
boxplot(min_to_subway, main = "Minutes to subway")
hist(building_age_yrs,xlab = "building age", main = "Building's age in years")

library(ggplot2)
a=ggplot(data=train, aes(x=bedrooms, y=rent)) + 
  geom_point() + 
  geom_smooth(method = lm, se = FALSE) + 
  labs(x = 'Number of bedrooms', y='Rent', 
       title = 'Bedrooms VS Rent')


b=ggplot(data=train, aes(x=size_sqft, y=rent)) + 
  geom_point() + 
  geom_smooth(method = lm, se = FALSE) + 
  labs(x = 'Size in sqft', y='Rent', 
       title = 'Unit VS rent')

grid.arrange(a,b, nrow=1)
```

powerTransform on y and EDA after transformation on y(rent)
```{r}
#install.packages("car")
library(car)
transform <- powerTransform(cbind(data1$rent))
summary(transform)
train$logRent=log(train$rent)
attach(train)
par(mfrow=c(3,2))
hist(logRent, breaks=10, main="logRent",col="#1793d1")
hist(bedrooms,breaks=10,xlab = "number of bedrooms", main="Number of Bedrooms")
hist(bathrooms, breaks=10,xlab = "number of bathrooms", main="Number of Bathrooms")
hist(size_sqft, breaks=10,xlab = "size", main="Size in sqft")
boxplot(min_to_subway, main = "Minutes to subway")
hist(building_age_yrs, xlab = "building age",main = "Building's age in years")

library(ggplot2)

grid.arrange(a,b, nrow=1)


a=ggplot(data=train, aes(x=bedrooms, y=logRent)) + 
  geom_point() + 
  geom_smooth(method = lm, se = FALSE) + 
  labs(x = 'Number of Bedrooms', y='logRent', 
       title = 'Bedrooms VS logRent')


b=ggplot(data=train, aes(x=size_sqft, y=logRent)) + 
  geom_point() + 
  geom_smooth(method = lm, se = FALSE) + 
  labs(x = 'Size in sqft', y='logRent', 
       title = 'Size in sqft VS logRent')

grid.arrange(a,b, nrow=1)                            
                          
```
Model 1
```{r}

library(kableExtra)
m1=lm(logRent~bedrooms+bathrooms+size_sqft+min_to_subway+building_age_yrs+floor+building_age_yrs+has_dishwasher+has_doorman+has_elevator+has_gym+has_roofdeck,data=train)
summary(m1)

```
Model 2, AIC, BIC, ajusted R square checking
```{r}
m2=lm(logRent~bedrooms+bathrooms+size_sqft+min_to_subway+building_age_yrs+floor+building_age_yrs,data=train)
summary(m2)

#SSres, adjusted R square, AIC, AICc, BIC check
summary(m2)

summary(m1)

#multicollinearity check
vif(m2)
```
Partial model, model 3 and paritial f test
```{r}
m3=lm(logRent~bedrooms+size_sqft+min_to_subway+building_age_yrs,data=train)
anova(m2,m3)


```
condition check for model 2 with plots
```{r}
r <- resid(m2)
#condition 1
plot(rent ~ fitted(m2), main="Y versus Y-hat", xlab="logRent-hat", ylab="logRent")
abline(a = 0, b = 1)
lines(lowess(rent ~ fitted(m2)), lty=2)
#condition 2
data2 = data.frame(train$rent, train$bedrooms, train$size_sqft)
pairs( data2 )
```
4 assumption checks of model 2 using residual plots, qqplots. cook distance plots
```{r}
par(mfrow=c(2,2))
##residual vs fitted plot
plot(m2,1)
##qqplot
plot(m2,2)
##residual vs X plots
plot(train$bedrooms, r, xlab="Number of bedrooms", ylab="Residuals", main="Residuals vs x1")
plot(train$bathrooms, r, xlab="Number of bathrooms", ylab="Residuals", main="Residuals vs x2")
par(mfrow=c(2,2))
plot(train$size_sqft, r, xlab="Size", ylab="Residuals", main="Residuals vs x3")
plot(train$min_to_subway, r, xlab="Minutes to subway", ylab="Residuals", main="Residuals vs x4")
plot(train$building_age_yrs, r, xlab="Building age", ylab="Residuals", main="Residuals vs x5")
plot(train$floor, r, xlab="Floor", ylab="Residuals", main="Residuals vs x6")
par(mfrow=c(2,2))
##cook distance plot
plot(m2,4)
```
outlier points, leverage points, influential points, and multicollinearity check
```{r}
#outlier points
r <- rstandard(m2)
out <- which(r > 2 | r < -2)
out

#leverage points
h <- hatvalues(m2)
threshold <- 2 * (length(m2$coefficients)/nrow(train))
w <- which(h > threshold)
train[w,]

#influential points
D <- cooks.distance(m2)
cutoff <- qf(0.5, length(m2$coefficients), nrow(train)-length(m2$coefficients), lower.tail=T)
which(D > cutoff)



```
fitting model(model 5) using test data and its plots, adjusted R square, AIC, BIC
```{r}
test$logRent=log(test$rent)
m5=lm(logRent~bedrooms+bathrooms+size_sqft+min_to_subway+building_age_yrs+floor+building_age_yrs,data=test)
summary(m5)


# Plots to check conditions
r <- resid(m5)
#condition 1
plot(test$rent ~ fitted(m5), main="Y versus Y-hat", xlab="logRent-hat", ylab="logRent")
abline(a = 0, b = 1)
lines(lowess(test$rent ~ fitted(m5)), lty=2)
#condition 2
data2 = data.frame(test$rent, test$bedrooms,test$size_sqft)
pairs( data2 )

#Plots to check assumptions
par(mfrow=c(2,2))
##residual vs fitted
plot(m5,1)
##qqplot
plot(m5,2)
##residual vs X
plot(test$bedrooms, r, xlab="Number of bedrooms", ylab="Residuals", main="Residuals vs x1")
plot(test$bathrooms, r, xlab="Number of bathrooms", ylab="Residuals", main="Residuals vs x2")
par(mfrow=c(2,2))
plot(test$size_sqft, r, xlab="Size", ylab="Residuals", main="Residuals vs x3")
plot(test$min_to_subway, r, xlab="Minutes to subway", ylab="Residuals", main="Residuals vs x4")
plot(test$building_age_yrs, r, xlab="Building age", ylab="Residuals", main="Residuals vs x5")
plot(test$floor, r, xlab="Floor", ylab="Residuals", main="Residuals vs x6")
par(mfrow=c(2,2))
#Extra:cook distance
plot(m5,4)

#SSres, adjusted R square, AIC, AICc, BIC check
summary(m2)
summary(m5)

```

